---
title: "Cyclistic Bike Share Case Study Analysis"
author: "Arjit Bhardwaj"
date: "2023-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(lubridate)
library(geosphere)
library(ggplot2)
library(ggmap)
library(forcats)
```

### **Introduction**

This is the capstone project as part of my Google Data Analytics Professional Certificate course. For the analysis, I will be using R programming language and RStudio IDE.

#### **Modus Operandi**

-   Ask
-   Prepare
-   Process
-   Analyse
-   Share
-   Act

### **Scenario**

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations.

### **Problem Statement**

-   How do annual members and casual riders use Cyclistic bikes differently?
-   Why would casual riders buy Cyclistic annual memberships?
-   How can Cyclistic use digital media to influence casual riders to become members?

The director of marketing and your manager Lily Moreno has assigned you the first question to answer: How do annual members and casual riders use Cyclistic bikes differently?

### **Ask**

#### **Key tasks**

-   Identify the business task
    -   The main objective is to build the best marketing strategies to turn casual bike riders into annual members by analyzing how the 'Casual' and 'Annual' customers use Cyclistic bike share differently.
-   Consider key stakeholders
    -   Cyclistic executive team
    -   Director of Marketing
    -   Marketing Analytics team

#### **Deliverable**

-   A clear statement of the business task
    -   Find the difference between casual and member riders.

### **Prepare**

I will be using Cyclistic's historical trip data to analyse and identify trends. The data has been made available by **Motivate International Inc.**

#### **Key Tasks**

-   Download data and store it appropriately.
    -   Data has been downloaded from [divvy-tripdata](https://divvy-tripdata.s3.amazonaws.com/index.html) and copies have been stored securely on my computer.
-   Identify how it's organised.
    -   All trip data is in comma-delimited (.CSV) format. Column names - "ride_id", "rideable_type", "started_at", "ended_at", "start_station_name", "start_station_id", "end_station_name","end_station_id", "start_lat", "start_lng", "end_lat", "end_lng", "member_casual" (Total 13 columns).
-   Sort and filter the data.
    -   For this analysis I'm going to use past 12 months data starting form May 2022 to April 2023.
-   Determine the credibility of the data.
    -   For the purpose of this case study, the datasets are appropriate and it will enable me to answer the business questions. But due to privacy policy, I cannot use rider's personally identification information, and this will prevent me from determining if a single rider taken several rides. All ride_id are unique in this datasets.

#### **Deliverable**

-   A description of all data sources used.
    -   Main source data provided by Cyclistic Company.

##### importing the packages to be used

```{r}
library(readr)
library(dplyr)
library(lubridate)
library(geosphere)
library(ggplot2)
library(ggmap)
```

##### import data in R Studio

```{r}
df1 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202205-divvy-tripdata.csv')
df2 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202206-divvy-tripdata.csv')
df3 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202207-divvy-tripdata.csv')
df4 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202208-divvy-tripdata.csv')
df5 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202209-divvy-tripdata.csv')
df6 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202210-divvy-tripdata.csv')
df7 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202211-divvy-tripdata.csv')
df8 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202212-divvy-tripdata.csv')
df9 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202301-divvy-tripdata.csv')
df10 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202302-divvy-tripdata.csv')
df11 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202303-divvy-tripdata.csv')
df12 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202304-divvy-tripdata.csv')
```

##### checking the consistency of data

```{r}
colnames(df1)
colnames(df2)
colnames(df3)
colnames(df4)
colnames(df5)
colnames(df6)
colnames(df7)
colnames(df8)
colnames(df9)
colnames(df10)
colnames(df11)
colnames(df12)
```

##### checking the structure of data

```{r}
str(df1)
str(df2)
str(df3)
str(df4)
str(df5)
str(df6)
str(df7)
str(df8)
str(df9)
str(df10)
str(df11)
str(df12)
```

##### merging 12 datasets into 1 dataset

```{r}
all_rides <- rbind(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12)
```

##### creating a file of the merged dataset

```{r}
write_csv(all_rides, '/home/arjit/Projects/Case Study Cyclistic/Data/all_rides.csv')
```

### **Process**

Cleaning and processing data for analysis

#### **Key tasks**

-   Check the data for errors
-   Choose your tools
-   Transform the data so you can work with it effectively
-   Document the cleaning process

#### **Deliverable**

-   Documentation of any cleaning or manipulating the data

##### Analyzing the the structure of attributes

```{r}
sapply(all_rides, class)
```

##### Number of columns, rows and total NA values

```{r}
ncol(all_rides)
nrow(all_rides)

sum(is.na(all_rides))
```

##### Removing all the empty cells

```{r}
all_rides <- janitor::remove_empty(all_rides, which=c("cols"))
all_rides <- janitor::remove_empty(all_rides, which=c("rows"))
```

##### Removing all the NA values

```{r}
all_rides <- na.omit(all_rides)
sum(is.na(all_rides))
```

##### Checking the number of rows left after removing NA

```{r}
nrow(all_rides)
```

##### Checking the dimension of the dataset

```{r}
dim(all_rides)
```

##### Calculating the ride time

```{r}
all_rides$minutes <- difftime(all_rides$ended_at, all_rides$started_at, units = c("min"))
all_rides$minutes <- as.numeric(as.character(all_rides$minutes))
all_rides$minutes <- round(all_rides$minutes, 1)
```

##### Checking the datatype of "minutes" column

```{r}
is.numeric(all_rides$minutes)
```

##### Adding date, month, day, year, day of week, hour columns

```{r}
all_rides$date <- as.Date(all_rides$started_at)
all_rides$month <- format(as.Date(all_rides$date), "%m")
all_rides$day <- format(as.Date(all_rides$date), "%d")
all_rides$year <- format(as.Date(all_rides$date), "%Y")
all_rides$day_of_week <- format(as.Date(all_rides$date), "%A")
all_rides$hour <- hour(all_rides$started_at)
```

##### Checking the modified dataset

```{r}
head(all_rides)
```

##### Adding the season column

```{r}
all_rides <- all_rides %>% mutate(season = case_when(month == "01" ~ "Winter",
                                                     month == "02" ~ "Winter", 
                                                     month == "03" ~ "Spring", 
                                                     month == "04" ~ "Spring", 
                                                     month == "05" ~ "Summer", 
                                                     month == "06" ~ "Summer", 
                                                     month == "07" ~ "Summer", 
                                                     month == "08" ~ "Summer", 
                                                     month == "09" ~ "Fall", 
                                                     month == "10" ~ "Fall", 
                                                     month == "11" ~ "Winter", 
                                                     month == "12" ~ "Winter"))
```

##### Mapping the time column to phases of a day

```{r}
all_rides <- all_rides %>% mutate(time_of_day = case_when(hour == "0" ~ "Night",
                                                          hour == "1" ~ "Night",
                                                          hour == "2" ~ "Night", 
                                                          hour == "3" ~ "Night", 
                                                          hour == "4" ~ "Night", 
                                                          hour == "5" ~ "Morning", 
                                                          hour == "6" ~ "Morning", 
                                                          hour == "7" ~ "Morning", 
                                                          hour == "8" ~ "Morning", 
                                                          hour == "9" ~ "Morning", 
                                                          hour == "10" ~ "Morning", 
                                                          hour == "11" ~ "Morning", 
                                                          hour == "12" ~ "Afternoon", 
                                                          hour == "13" ~ "Afternoon", 
                                                          hour == "14" ~ "Afternoon", 
                                                          hour == "15" ~ "Afternoon", 
                                                          hour == "16" ~ "Afternoon", 
                                                          hour == "17" ~ "Afternoon", 
                                                          hour == "18" ~ "Evening", 
                                                          hour == "19" ~ "Evening", 
                                                          hour == "20" ~ "Evening", 
                                                          hour == "21" ~ "Evening", 
                                                          hour == "22" ~ "Evening", 
                                                          hour == "23" ~ "Evening"))
```

##### Mapping the months to their names

```{r}
all_rides <- all_rides %>% mutate(month = case_when(month == "01" ~ "January", 
                                                    month == "02" ~ "February", 
                                                    month == "03" ~ "March", 
                                                    month == "04" ~ "April", 
                                                    month == "05" ~ "May", 
                                                    month == "06" ~ "June", 
                                                    month == "07" ~ "July", 
                                                    month == "08" ~ "August", 
                                                    month == "09" ~ "September", 
                                                    month == "10" ~ "October", 
                                                    month == "11" ~ "November", 
                                                    month == "12" ~ "December"))
```

##### Verifying the results

```{r}
tail(all_rides)
```

### **Analyze**

Now all the required information are in one place and ready for exploration.

#### **Key tasks**

-   Aggregate your data so it's useful and accessible.
-   Organise and format your data.
-   Perform calculations.
-   Identify trends and relationships.

#### **Deliverable**

-   A summary of analysis.

##### Checking the unique types of ride available

```{r}
unique(all_rides$rideable_type)
```

##### Checking the unique values in member_casual

```{r}
unique(all_rides$member_casual)
```

##### Verifying if there are any NA in dataset

```{r}
sum(is.na(all_rides))
```

##### Checking if there are any duplicates

```{r}
#distinct(all_rides)
```

```{r}
dim(all_rides)
```

```{r}
write_csv(all_rides, '/home/arjit/Projects/Case Study Cyclistic/Data/all_rides_modified.csv')
```

```{r}
df <- read_csv('/home/arjit/Projects/Case Study Cyclistic/Data/all_rides_modified.csv')
```

```{r}
head(df)
```

```{r}
dim(df)
```

##### Checking the structure of dataset

```{r}
str(df)
```

##### Analyzing the summary of the dataset

```{r}
summary(df)
```

##### Calculating the ride length in kilometers

```{r}
df$ride_length <- distGeo(matrix(c(df$start_lng, df$start_lat), ncol = 2), matrix(c(df$end_lng, df$end_lat), ncol = 2))

df$ride_length <- df$ride_length/1000
```

##### Removing the rows with ride length as negative or zero

```{r}
df <- df[!(df$ride_length <= 0),]
```

##### Removing the NA in the dataset

```{r}
sum(is.na(df))

df <- na.omit(df)
```

#### Doing descriptive analysis, studying

##### avg_ride_length -\> straight average(total ride length / total rides)

##### median_length -\> midpoint number of ride length

##### max_ride_length -\> longest ride

##### min_ride_length -\> shortest ride

```{r}
df %>% summarise(avg_ride_length = mean(ride_length), median_length = median(ride_length), 
                 max_ride_length = max(ride_length), min_ride_length = min(ride_length))
```

#### **Compare members and casual riders**

-   Member vs casual riders difference depending on the total rides taken

```{r}
df %>% 
    group_by(member_casual) %>% 
    summarise(ride_count = length(ride_id), ride_percentage = (length(ride_id) / nrow(df)) * 100)

ggplot(df, aes(x = member_casual, fill=member_casual)) +
    geom_bar() +
    labs(x="Casuals vs Members", y="Number Of Rides", title= "Casuals vs Members distribution")
```

We can see on the Casuals vs Members distribution chart, members possessing \~ 61%, and casual riders have \~ 39% of the dataset. So it is clearly visible that in the whole year 2022 members used ride share 22% more than casual riders.

#### Comparison between Member Casual riders depending on ride length

```{r}
df %>%
  group_by(member_casual) %>% 
  summarise(avg_ride_length = mean(ride_length), median_length = median(ride_length), 
            max_ride_length = max(ride_length), min_ride_length = min(ride_length))
```

From the above table we can conclude that casual riders took bike for longer rides than members, as the average trip duration / average ride length of member riders is lower than the average trip duration / average ride length of casual riders.

#### Fixing the names of days in a week

```{r}
df$day_of_week <- ordered(df$day_of_week, 
                                    levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
df %>% 
  group_by(member_casual, day_of_week) %>%  
  summarise(number_of_rides = n()
  ,avg_ride_length = mean(ride_length),.groups="drop") %>% 
  arrange(member_casual, day_of_week)
```

#### Analyzing the total rides by members and casual riders in a particular day in a week

```{r}
df %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n(), .groups="drop") %>% 
  arrange(member_casual, day_of_week)  %>% 
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  labs(title ="Total rides by Members and Casual riders Vs. Day of the week") +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))


```

From the above graph, we can conclude that members took consistent rides throughout the week with least on Sunday. For the casual riders, the most rides were taken on weekends.

#### Analyzing the average ride time of members casual riders on a particular day of a week

```{r}
df %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(average_ride_length = mean(ride_length), .groups="drop") %>%
  ggplot(aes(x = day_of_week, y = average_ride_length, fill = member_casual)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) + 
  labs(title ="Average ride time by Members and Casual riders Vs. Day of the week")
```

The average ride length for members are comparatively less than that of casual riders. Also it can be seen that weekend average ride length is much higher for casual riders along with total rides. So both of this facts can be correlated for casual riders. For members average ride length is about the same throughout the week

#### Grouping the rides according to member casual and months of a year

```{r}
df %>% 
  group_by(member_casual, month) %>%  
  summarise(number_of_rides = n(), average_ride_length = mean(ride_length), .groups="drop") %>% 
  arrange(member_casual, month)
```

#### Analyzing the total rides by members and casual riders in a month

```{r}
df$month <- ordered(df$month, 
                            levels=c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))
df %>%  
  group_by(member_casual, month) %>% 
  summarise(number_of_rides = n(),.groups="drop") %>% 
  arrange(member_casual, month)  %>% 
  ggplot(aes(x = month, y = number_of_rides, fill = member_casual)) +
  labs(title ="Total rides by Members and Casual riders Vs. Month", x = "Month", y= "Number Of Rides") +
  theme(axis.text.x = element_text(angle = 45)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

The months June, July, August and September are the most busy time of the year among both members and casual riders. It is possible due to winter there is a significant drop in total rides in the months of November, December, January and February for both type of customers. But we can see that member's total rides are higher than casual riders throughout the year except from June, July and August.

#### Analyzing the ride length by members casual riders in a month

```{r}
df %>%  
  group_by(member_casual, month) %>% 
  summarise(average_ride_length = mean(ride_length),.groups="drop") %>%
  ggplot(aes(x = month, y = average_ride_length, fill = member_casual)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) + 
  labs(title ="Average ride length by Members and Casual riders Vs. Month") +
  theme(axis.text.x = element_text(angle = 30))
```

Average ride length of members is about the same throughout the year. While casual riders average ride length is greater than the members throughout the year.

#### Analyzing the average distance traveled by member casual rider

```{r}
df %>% 
  group_by(member_casual) %>%
  summarise(avg_ride_length = mean(ride_length)) %>%
  ggplot() + 
  geom_col(mapping= aes(x= member_casual,y= avg_ride_length,fill=member_casual), show.legend = FALSE)+
  labs(title = "Mean travel distance by Members and Casual riders", x="Member and Casual riders", y="Average distance In Km")
```

From the above chart we can see that both riders travel about the same average distance. This similarity could be possible due to that member take (same ride time) rides throughout the week, but casual riders took rides mostly in weekends with higher ride time.

#### Analysis and visualize the bike demand by hour in a day

```{r}
df %>%
    ggplot(aes(hour, fill= member_casual)) +
    labs(x="Hour of the day", title="Cyclistic's Bike demand by hour in a day") +
    geom_bar()
```

From the above chart we can see more members between 7am and 11am and more casual riders between 3pm and 12am. Also there is bigger volume rise in the afternoon for both type of riders. This information needs to be checked on day basis.

#### Analysis and visualize the bikes demand by hour in a day of a week

```{r}
df %>%
    ggplot(aes(hour, fill=member_casual)) +
    geom_bar() +
    labs(x="Hour of the day", title="Cyclistic's bike demand per hour by day of the week") +
    facet_wrap(~ day_of_week)
```

There is a lot of difference between the weekdays and weekends. There is a big increase of volume in the weekdays between 7am to 10am and another volume increase from 5pm to 7pm. We can hypothesize that members use the bikes as daily routine like going to work (same behavior throughout the weekdays) and go back from work (5pm - 7pm). Weekends are completely different for members and casual riders, Friday, Saturday and Sunday there is huge peak in volume for casual riders, from this we can hypothesize that casual riders mostly use bike share for leisure activity in the weekends.

#### Analyzing the preferred bike by member casual riders

```{r}
df %>%
    group_by(rideable_type) %>% 
    summarise(count = length(ride_id))

ggplot(df, aes(x=rideable_type, fill=member_casual)) +
    labs(x="Rideable type", title="Rideable type Vs. total rides by Members and casual riders") +
    geom_bar()
```

From the above viz we can see that members mostly use classic bikes, followed by electric bikes. Docked bikes mostly used by casual riders. Electric bikes are more favored by members.

#### Analyzing and visualizing the top 5 starting stations for casual riders

```{r}
df %>% 
  filter(!(is.na(start_station_name))) %>% 
  filter(member_casual == "casual") %>% 
  group_by(start_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5)
```

```{r}
df %>% 
  filter(!(is.na(start_station_name))) %>% 
  filter(member_casual == "casual") %>% 
  group_by(start_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5) %>% 
  mutate(start_station_name= fct_reorder(start_station_name, count)) %>% 
  ggplot(aes(x=start_station_name, y=count, fill=count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x="Number of Rides", y="Start Station Name", title="Top 5 starting stations for casual riders")
```

#### Analyzing and visualizing the top 5 ending stations for casual riders

```{r}
df %>% 
  filter(!(is.na(end_station_name))) %>% 
  filter(member_casual == "casual") %>% 
  group_by(end_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5)
```

```{r}
df %>% 
  filter(!(is.na(end_station_name))) %>% 
  filter(member_casual == "casual") %>% 
  group_by(end_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5) %>% 
  mutate(end_station_name= fct_reorder(end_station_name, count)) %>% 
  ggplot(aes(x=end_station_name, y=count, fill=count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x="Number of Rides", y="End Station Name", title="Top 5 ending stations for casual riders")
```

#### Analyzing and visualizing the top 5 starting stations for member riders

```{r}
df %>% 
  filter(!(is.na(start_station_name))) %>% 
  filter(member_casual == "member") %>% 
  group_by(start_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5)
```

```{r}
df %>% 
  filter(!(is.na(start_station_name))) %>% 
  filter(member_casual == "member") %>% 
  group_by(start_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5) %>% 
  mutate(start_station_name= fct_reorder(start_station_name, count)) %>% 
  ggplot(aes(x=start_station_name, y=count, fill=count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x="Number of Rides", y="Start Station Name", title="Top 5 starting stations for member riders")
```

#### Analyzing and visualizing the top 5 ending stations for casual riders

```{r}
df %>% 
  filter(!(is.na(end_station_name))) %>% 
  filter(member_casual == "member") %>% 
  group_by(end_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5)
```

```{r}
df %>% 
  filter(!(is.na(end_station_name))) %>% 
  filter(member_casual == "member") %>% 
  group_by(end_station_name) %>% 
  summarize(count=n()) %>% 
  arrange(-count) %>% 
  top_n(5) %>% 
  mutate(end_station_name= fct_reorder(end_station_name, count)) %>% 
  ggplot(aes(x=end_station_name, y=count, fill=count)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x="Number of Rides", y="End Station Name", title="Top 5 ending stations for member riders")
```

#### **Analyze and visualize the dataset on coordinate basis**

```{r}
coordinates_df <- df %>% 
filter(start_lng != end_lng & start_lat != end_lat) %>%
group_by(start_lng, start_lat, end_lng, end_lat, member_casual, rideable_type) %>%
summarise(total_rides = n(),.groups="drop") %>%
filter(total_rides > 200)
```

```{r}
casual_riders <- coordinates_df %>% filter(member_casual == "casual")
member_riders <- coordinates_df %>% filter(member_casual == "member")
```

#### Setting up the ggmap for Chicago

```{r}
chicago <- c(left = -87.700424, bottom = 41.790769, right = -87.554855, top = 41.990119)

chicago_map <- get_stamenmap(bbox = chicago, zoom = 12, maptype = "terrain")
```

#### Visualization on the map

```{r}
ggmap(chicago_map,darken = c(0.1, "white")) +
   geom_point(casual_riders, mapping = aes(x = start_lng, y = start_lat, color=rideable_type), size = 2) +
   coord_fixed(0.8) +
   labs(title = "Most used routes by Casual riders",x=NULL,y=NULL) +
   theme(legend.position="none")


ggmap(chicago_map,darken = c(0.1, "white")) +
    geom_point(member_riders, mapping = aes(x = start_lng, y = start_lat, color=rideable_type), size = 2) +  
    coord_fixed(0.8) +
    labs(title = "Most used routes by Member riders",x=NULL,y=NULL) +
    theme(legend.position="none")
```

We can clearly see the casual rides are mostly located around the center of the town (or the bay area), with all their trips located around that area points towards their bike usage pattern, which is for leisure, probably tourist or sightseeing related rides.

Members are mostly use bike all over the city including main city area and outside main center. This can be hypothesize as they travel for work purpose.

### **Share**

This phase will be done by presentation, but here we can use RMarkdown to share our analysis and visualizations.

#### **Key tasks**

-   Determine the best way to share your findings.
-   Create effective data visualizations.
-   Present your findings.
-   Ensure your work is accessible.

#### **Deliverable**

-   Support visualizations and key findings.

#### **Main insights and conclusions**

-   Members holds the biggest proportion of the total rides, \~22% bigger than casual riders.
-   In all months we have more members than casual riders.
-   For casual riders the biggest volume of data is on the weekend.
-   There is a bigger volume of bikers in the afternoon. This could be possible that member use bikes for work purpose, this information can be backed by their bike usage in colder months, where there is significant drop in casual members in those months.

#### **How members differs from casuals:**

-   Members have the bigger volume of data, except on Saturday and Sunday. On the weekend, casuals riders have the most rides.
-   Casuals riders have more ride length (ride duration) than members. Average ride time of member are mostly same slight increase in end of week.
-   We have more members during the morning, mainly between 7am and 10am. And more casuals between 3pm and 12am.
-   Members have a bigger preference for classic bikes, followed by electric bike.
-   Members have a more fixed use for bikes for routine activities. Where as casual rider's usage is different, mostly all activity in the weekend.
-   Casual member spend time near the center of the city or the bay area, where as member are scattered throughout the city.

### **Act**

Act phase will be done by the Cyclistic's executive team, Director of Marketing (Lily Moreno), Marketing Analytics team on the basis of my analysis. (Data-driven decision making)

#### **Deliverable**

-   Top three recommendations:
    -   Offer a weekend-only membership at a different price point than the full annual membership.

    -   Coupons and discounts could be handed out along with the annual subscription / weekend-only membership for the usage of electric bikes targeting casual riders. Also increasing the number of electric bike while reducing classic bikes if electric bike costs more for the pass, this can be beneficial for the company. (As electric bike are already in trend and usage is good as per member and ride type data.

    -   Create marketing campaigns which can be sent via email, or advertisement in the docking stations explaining why annual member is beneficial. Campaigns should be placed at the peak months of the year.
-   **Additional Recommendations**:
    -   Raise awareness of the ease of use, affordability, accessibility and green options.
    -   Focus on dense and walk able areas, partnering with school and companies.

#### **Conclusion**

Thank you for your time and interest to review my capstone project! This project helped me to walk through the data analysis process from start to finish using real-world data and business questions. To learn from the others code too, I have referred the analysis done by Elaine Sanchez Wilber and Sayantan Bagchi. I'm truly excited and look forward to growing in the field of data analysis.
