---
title: "Cyclistic Bike Share Case Study Analysis"
author: "Arjit Bhardwaj"
date: "2023-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(lubridate)
library(geosphere)
library(ggplot2)
library(ggmap)
```

### **Introduction**

This is the capstone project as part of my Google Data Analytics Professional Certificate course. For the analysis, I will be using R programming language and RStudio IDE.

#### **Modus Operandi**

-   Ask
-   Prepare
-   Process
-   Analyse
-   Share
-   Act

### **Scenario**

You are a junior data analyst working in the marketing analyst team at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore, your team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, your team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve your recommendations, so they must be backed up with compelling data insights and professional data visualizations.

### **Problem Statement**

-   How do annual members and casual riders use Cyclistic bikes differently?
-   Why would casual riders buy Cyclistic annual memberships?
-   How can Cyclistic use digital media to influence casual riders to become members?

The director of marketing and your manager Lily Moreno has assigned you the first question to answer: How do annual members and casual riders use Cyclistic bikes differently?

### **Ask**

#### **Key tasks**

-   Identify the business task
    -   The main objective is to build the best marketing strategies to turn casual bike riders into annual members by analyzing how the 'Casual' and 'Annual' customers use Cyclistic bike share differently.
-   Consider key stakeholders
    -   Cyclistic executive team
    -   Director of Marketing
    -   Marketing Analytics team

#### **Deliverable**

-   A clear statement of the business task
    -   Find the difference between casual and member riders.

### **Prepare**

I will be using Cyclistic's historical trip data to analyse and identify trends. The data has been made available by **Motivate International Inc.**

#### **Key Tasks**

-   Download data and store it appropriately.
    -   Data has been downloaded and copies have been stored securely on my computer.
-   Identify how it's organised.
    -   All trip data is in comma-delimited (.CSV) format. Column names - "ride_id", "rideable_type", "started_at", "ended_at", "start_station_name", "start_station_id", "end_station_name","end_station_id", "start_lat", "start_lng", "end_lat", "end_lng", "member_casual" (Total 13 columns).
-   Sort and filter the data.
    -   For this analysis I'm going to use past 12 months data starting form May 2022 to April 2023.
-   Determine the credibility of the data.
    -   For the purpose of this case study, the datasets are appropriate and it will enable me to answer the business questions. But due to privacy policy, I cannot use rider's personally identification information, and this will prevent me from determining if a single rider taken several rides. All ride_id are unique in this datasets.

#### **Deliverable**

-   A description of all data sources used.
    -   Main source data provided by Cyclistic Company.

##### importing the packages to be used

```{r}
library(readr)
library(dplyr)
library(lubridate)
library(geosphere)
library(ggplot2)
library(ggmap)
```

##### import data in R Studio

```{r}
df1 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202205-divvy-tripdata.csv')
df2 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202206-divvy-tripdata.csv')
df3 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202207-divvy-tripdata.csv')
df4 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202208-divvy-tripdata.csv')
df5 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202209-divvy-tripdata.csv')
df6 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202210-divvy-tripdata.csv')
df7 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202211-divvy-tripdata.csv')
df8 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202212-divvy-tripdata.csv')
df9 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202301-divvy-tripdata.csv')
df10 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202302-divvy-tripdata.csv')
df11 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202303-divvy-tripdata.csv')
df12 <- read.csv('/home/arjit/Projects/Case Study Cyclistic/Data/202304-divvy-tripdata.csv')
```

##### checking the consistency of data

```{r}
colnames(df1)
colnames(df2)
colnames(df3)
colnames(df4)
colnames(df5)
colnames(df6)
colnames(df7)
colnames(df8)
colnames(df9)
colnames(df10)
colnames(df11)
colnames(df12)
```

##### checking the structure of data

```{r}
str(df1)
str(df2)
str(df3)
str(df4)
str(df5)
str(df6)
str(df7)
str(df8)
str(df9)
str(df10)
str(df11)
str(df12)
```

##### merging 12 datasets into 1 dataset

```{r}
all_rides <- rbind(df1, df2, df3, df4, df5, df6, df7, df8, df9, df10, df11, df12)
```

##### creating a file of the merged dataset

```{r}
write_csv(all_rides, '/home/arjit/Projects/Case Study Cyclistic/Data/all_rides.csv')
```

```{r}
sapply(all_rides, class)
```

```{r}
ncol(all_rides)
nrow(all_rides)

sum(is.na(all_rides))
```

```{r}
all_rides <- janitor::remove_empty(all_rides, which=c("cols"))
all_rides <- janitor::remove_empty(all_rides, which=c("rows"))
```

```{r}
all_rides <- na.omit(all_rides)
sum(is.na(all_rides))
```

```{r}
nrow(all_rides)
```

```{r}
dim(all_rides)
```

```{r}
all_rides$minutes <- difftime(all_rides$ended_at, all_rides$started_at, units = c("min"))
all_rides$minutes <- as.numeric(as.character(all_rides$minutes))
all_rides$minutes <- round(all_rides$minutes, 1)
```

```{r}
is.numeric(all_rides$minutes)
```

```{r}
all_rides$date <- as.Date(all_rides$started_at)
all_rides$month <- format(as.Date(all_rides$date), "%m")
all_rides$day <- format(as.Date(all_rides$date), "%d")
all_rides$year <- format(as.Date(all_rides$date), "%Y")
all_rides$day_of_week <- format(as.Date(all_rides$date), "%A")
all_rides$hour <- hour(all_rides$started_at)
```

```{r}
head(all_rides)
```

```{r}
all_rides <- all_rides %>% mutate(season = case_when(month == "01" ~ "Winter",
                                                     month == "02" ~ "Winter", 
                                                     month == "03" ~ "Spring", 
                                                     month == "04" ~ "Spring", 
                                                     month == "05" ~ "Summer", 
                                                     month == "06" ~ "Summer", 
                                                     month == "07" ~ "Summer", 
                                                     month == "08" ~ "Summer", 
                                                     month == "09" ~ "Fall", 
                                                     month == "10" ~ "Fall", 
                                                     month == "11" ~ "Winter", 
                                                     month == "12" ~ "Winter"))
```

```{r}
all_rides <- all_rides %>% mutate(time_of_day = case_when(hour == "0" ~ "Night",
                                                          hour == "1" ~ "Night",
                                                          hour == "2" ~ "Night", 
                                                          hour == "3" ~ "Night", 
                                                          hour == "4" ~ "Night", 
                                                          hour == "5" ~ "Morning", 
                                                          hour == "6" ~ "Morning", 
                                                          hour == "7" ~ "Morning", 
                                                          hour == "8" ~ "Morning", 
                                                          hour == "9" ~ "Morning", 
                                                          hour == "10" ~ "Morning", 
                                                          hour == "11" ~ "Morning", 
                                                          hour == "12" ~ "Afternoon", 
                                                          hour == "13" ~ "Afternoon", 
                                                          hour == "14" ~ "Afternoon", 
                                                          hour == "15" ~ "Afternoon", 
                                                          hour == "16" ~ "Afternoon", 
                                                          hour == "17" ~ "Afternoon", 
                                                          hour == "18" ~ "Evening", 
                                                          hour == "19" ~ "Evening", 
                                                          hour == "20" ~ "Evening", 
                                                          hour == "21" ~ "Evening", 
                                                          hour == "22" ~ "Evening", 
                                                          hour == "23" ~ "Evening"))
```

```{r}
all_rides <- all_rides %>% mutate(month = case_when(month == "01" ~ "January", 
                                                    month == "02" ~ "February", 
                                                    month == "03" ~ "March", 
                                                    month == "04" ~ "April", 
                                                    month == "05" ~ "May", 
                                                    month == "06" ~ "June", 
                                                    month == "07" ~ "July", 
                                                    month == "08" ~ "August", 
                                                    month == "09" ~ "September", 
                                                    month == "10" ~ "October", 
                                                    month == "11" ~ "November", 
                                                    month == "12" ~ "December"))
```

```{r}
tail(all_rides)
```

```{r}
unique(all_rides$rideable_type)
```

```{r}
unique(all_rides$member_casual)
```

```{r}
sum(is.na(all_rides))
```

```{r}
distinct(all_rides)
```

```{r}
dim(all_rides)
```

```{r}
write_csv(all_rides, '/home/arjit/Projects/Case Study Cyclistic/Data/all_rides_modified.csv')
```

```{r}
df <- read_csv('/home/arjit/Projects/Case Study Cyclistic/Data/all_rides_modified.csv')
```

```{r}
head(df)
```

```{r}
dim(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```

```{r}
df$ride_length <- distGeo(matrix(c(df$start_lng, df$start_lat), ncol = 2), matrix(c(df$end_lng, df$end_lat), ncol = 2))

df$ride_length <- df$ride_length/1000
```

```{r}
df <- df[!(df$ride_length <= 0),]
```

```{r}
sum(is.na(df))

df <- na.omit(df)
```

```{r}
str(df)
```

```{r}
df %>% summarise(avg_ride_length = mean(ride_length), median_length = median(ride_length), 
                 max_ride_length = max(ride_length), min_ride_length = min(ride_length))
```

```{r}
df %>% 
    group_by(member_casual) %>% 
    summarise(ride_count = length(ride_id), ride_percentage = (length(ride_id) / nrow(df)) * 100)

ggplot(df, aes(x = member_casual, fill=member_casual)) +
    geom_bar() +
    labs(x="Casuals vs Members", y="Number Of Rides", title= "Casuals vs Members distribution")
```

```{r}
df %>%
  group_by(member_casual) %>% 
  summarise(avg_ride_length = mean(ride_length), median_length = median(ride_length), 
            max_ride_length = max(ride_length), min_ride_length = min(ride_length))
```

```{r}
df %>% 
  group_by(member_casual, day_of_week) %>%  
  summarise(number_of_rides = n()
  ,avg_ride_length = mean(ride_length),.groups="drop") %>% 
  arrange(member_casual, day_of_week)
```

```{r}
df %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(number_of_rides = n(), .groups="drop") %>% 
  arrange(member_casual, day_of_week)  %>% 
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  labs(title ="Total rides by Members and Casual riders Vs. Day of the week") +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))


```

```{r}
df %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(average_ride_length = mean(ride_length), .groups="drop") %>%
  ggplot(aes(x = day_of_week, y = average_ride_length, fill = member_casual)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) + 
  labs(title ="Average ride time by Members and Casual riders Vs. Day of the week")
```

```{r}
df %>% 
  group_by(member_casual, month) %>%  
  summarise(number_of_rides = n(), average_ride_length = mean(ride_length), .groups="drop") %>% 
  arrange(member_casual, month)
```

```{r}
df %>%  
  group_by(member_casual, month) %>% 
  summarise(number_of_rides = n(),.groups="drop") %>% 
  arrange(member_casual, month)  %>% 
  ggplot(aes(x = month, y = number_of_rides, fill = member_casual)) +
  labs(title ="Total rides by Members and Casual riders Vs. Month", x = "Month", y= "Number Of Rides") +
  theme(axis.text.x = element_text(angle = 45)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

```{r}
df %>%  
  group_by(member_casual, month) %>% 
  summarise(average_ride_length = mean(ride_length),.groups="drop") %>%
  ggplot(aes(x = month, y = average_ride_length, fill = member_casual)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) + 
  labs(title ="Average ride length by Members and Casual riders Vs. Month") +
  theme(axis.text.x = element_text(angle = 30))
```

```{r}
df %>% 
  group_by(member_casual) %>%
  summarise(avg_ride_length = mean(ride_length)) %>%
  ggplot() + 
  geom_col(mapping= aes(x= member_casual,y= avg_ride_length,fill=member_casual), show.legend = FALSE)+
  labs(title = "Mean travel distance by Members and Casual riders", x="Member and Casual riders", y="Average distance In Km")
```

```{r}
df %>%
    ggplot(aes(hour, fill= member_casual)) +
    labs(x="Hour of the day", title="Cyclistic's Bike demand by hour in a day") +
    geom_bar()
```

```{r}
df %>%
    ggplot(aes(hour, fill=member_casual)) +
    geom_bar() +
    labs(x="Hour of the day", title="Cyclistic's bike demand per hour by day of the week") +
    facet_wrap(~ day_of_week)
```

```{r}
df %>%
    group_by(rideable_type) %>% 
    summarise(count = length(ride_id))

ggplot(df, aes(x=rideable_type, fill=member_casual)) +
    labs(x="Rideable type", title="Rideable type Vs. total rides by Members and casual riders") +
    geom_bar()
```

```{r}
coordinates_df <- df %>% 
filter(start_lng != end_lng & start_lat != end_lat) %>%
group_by(start_lng, start_lat, end_lng, end_lat, member_casual, rideable_type) %>%
summarise(total_rides = n(),.groups="drop") %>%
filter(total_rides > 200)
```

```{r}
casual_riders <- coordinates_df %>% filter(member_casual == "casual")
member_riders <- coordinates_df %>% filter(member_casual == "member")
```

```{r}
chicago <- c(left = -87.700424, bottom = 41.790769, right = -87.554855, top = 41.990119)

chicago_map <- get_stamenmap(bbox = chicago, zoom = 12, maptype = "terrain")
```

```{r}
ggmap(chicago_map,darken = c(0.1, "white")) +
   geom_point(casual_riders, mapping = aes(x = start_lng, y = start_lat, color=rideable_type), size = 2) +
   coord_fixed(0.8) +
   labs(title = "Most used routes by Casual riders",x=NULL,y=NULL) +
   theme(legend.position="none")


ggmap(chicago_map,darken = c(0.1, "white")) +
    geom_point(member_riders, mapping = aes(x = start_lng, y = start_lat, color=rideable_type), size = 2) +  
    coord_fixed(0.8) +
    labs(title = "Most used routes by Member riders",x=NULL,y=NULL) +
    theme(legend.position="none")
```
